---
title: "Network Exploration"
output: html_notebook
---

## Basic Setup and Get the Data

```{r setup, message=FALSE}

# Load libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse,         # Includes dplyr, ggplot2, and others; very key!
               visNetwork,        # For actually creating the visualization
               RColorBrewer,      # For generating a color palette
               googlesheets4)      # For retrieving the underlying data from Google Sheets

googlesheets_id <- "14QZbaKMNqGAGQpOPRZDZeHdb9H8kb19Vbv1VBItvpPs"

# Assuming the Google Sheet is public, then we can turn off the need to auth
gs4_deauth()

# Get the data
links_raw <- read_sheet(googlesheets_id, sheet="Data Flows")
nodes_raw <- read_sheet(googlesheets_id, sheet="Systems")

```

## Data Munging / Transformation

Convert to network graphing ready objects. This is one set of edges (the data feeds) and one set of nodes (the data sources). The "links" are the "edges" (the lines connecting the different data sources). This part of the script, essentially, renames the columns in the data table to match the visNetwork() function's 'edges' argument requirements later in the script. Of course, you could just name things that way in the underlying Google Sheet, but underlying Google Sheets are rarely designed with that much friendliness.

```{r data-munging, message=FALSE}

links <- links_raw %>% 
  # Select the columns that are actually going to be used
  select(from = From,
         to = To)
# The "nodes" are the actual data sources (the circles)
nodes <- nodes_raw %>% 
  select(id = System,
         data_type = `Data Type`,
         owner = Owner)

####################
# We'll size the nodes based on the number of data flows to/from them. The sizing
# could be anything (it could be something to do with the size of the actual system, 
# for instance), so this code can be tweaked accordingly.
# Get the total feeds going *to* each data source
nodes_to_links <- nodes %>%
  inner_join(links, by = c(id = "to")) %>% 
  group_by(id) %>% 
  summarise(total_links_to = n())

# Get the total feeds going *from* each data source
nodes_from_links <- nodes %>%
  inner_join(links, by = c(id = "from")) %>% 
  group_by(id) %>% 
  summarise(total_links_from = n())

# Combine those together -- adding *to* and *from*
nodes_link_counts <- nodes_to_links %>% 
  full_join(nodes_from_links) %>% 
  mutate(total_links_to = ifelse(is.na(total_links_to), 
                                 0, total_links_to)) %>%       # Replace NAs with 0s for future math
  mutate(total_links_from = ifelse(is.na(total_links_from), 
                                   0, total_links_from)) %>%   # Replace NAs with 0s for future math
  mutate(size = total_links_to + total_links_from) %>% 
  select(id, size)

# Join those back to the main nodes data
nodes <- nodes %>% 
  left_join(nodes_link_counts) %>% 
  filter(!is.na(size)) %>%             # Remove orphaned nodes
  mutate(size = 5 + size * 3)          # Scale the circle sizes (requires some fiddling)

# Remove unused data frames
rm(nodes_to_links, nodes_from_links, nodes_link_counts)
```

## Visualization

```{r visualization, message=FALSE}
# Add some more options
nodes$shape <- "dot"             # Just use circles
nodes$shadow <- FALSE            # Nodes will drop shadow
nodes$title <- nodes$id          # Node title
nodes$label <- nodes$id          # Node label
nodes$borderWidth <- 0.5         # Node border width
nodes$group <- nodes$data_type   # Use data type for the legend
nodes$color.border <- "gray80"
links$color <- "#BBBBBB"    # line color  
links$arrows <- "middle"    # arrows: 'from', 'to', or 'middle'
links$smooth <- TRUE        # should the edges be curved?
links$shadow <- FALSE       # edge shadow

# Generate the plot. The visGroups are a hack here. There's some risk that they won't
# wind up being quite right if a new category gets introduced.
viz_output <- visNetwork(nodes, links, width = "1024px", height = "768px") %>%
  visOptions(highlightNearest = TRUE,
             selectedBy = list(variable = "owner", main="Select System Owner")) %>% 
      
visLayout(randomSeed = 5003, improvedLayout = TRUE) %>% 
      visLegend(position = "right")
    
# Output the visualization
viz_output
```